---
title: "data_viz_steven"
author: "Steven Cognac"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## General use
library(tidyverse)
library(janitor)
library(here)

## specifically for data viz
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

library(sf)


```
## Get the lizards data
```{r}

lizards <- read_csv(here("data_tidy", "lizards.csv"))
```
## make histogram of lizard weight

```{r}
## if you are changing aesthics based on a constant, does not need to be in aesthetics

lizard_hist <- ggplot(data = lizards, aes(weight)) +
  geom_histogram(fill = "orange",
                 color = "blue",
                 size = 0.5,
                 linetype = "dotdash") +
  labs(x = "Weight",
       y = "Counts")

lizard_hist


## can also do if you want to stack different types of plots
lizard_hist2 <- ggplot() +
  geom_histogram(data = lizards, aes(weight)) +
  labs(x = "Weight",
       y = "Counts")



```

## scatterplot with total_length on x-axis and weight on y-axis
```{r}

## each shape from 0 - 25 represents different dot graphics
# alpha - transparency

lizard_scatter <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 22,
             fill = "yellow", # fill color
             color = "red", # color of shape outline
             size = 1.5, # size of dots
             alpha = 0.2) # transparency


lizard_scatter
```

## Make a bubble plot where color of points changes based on common_name and size of points varies based on total_length. total_length x axis, weight y axis
```{r}
lizards_bubble <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name,
                 size = total_length),
             shape = 22,
             fill = "black") +
  theme_light() +
  facet_grid()

lizards_bubble

```

## redo of above
```{r}

lizards_bubble2 <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_wrap(~common_name, ncol = 3)

lizards_bubble2
```
## 
```{r}
lizards_bubble3 <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_grid(sex ~ tail)

lizards_bubble3

```
## plot lizard counts by common_name
```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n()) %>% 
  mutate(common_name = fct_reorder(common_name, count))

lizard_counts


# same thing
lizard_count2 <- lizards %>% count(common_name)
lizard_count2

lizard_counts_cn_tail <- lizards %>% count(common_name, tail)
lizard_counts_cn_tail



```

## plot lizard counts
```{r}
# x and y axis switched to make it more readable.
# could also use the + coord_flip() function
# fct_reorder() lets you reorder some observation by a specified class

lizard_count_plot <- ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, count), x = count)) +
  geom_col(fill = "blue") +
  labs(x = "Lizard counts",
       y = "Common Name")

lizard_count_plot
```

## Axis scale breaks, limits, & labels

### scatterplot: total_length (x) versus weight (y)
```{r}
weight_length_plot <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(50, 100, 150, 200, 300, 400, 500),
                     limits = c(0, 600),
                     expand = c(0, 100)) +
  scale_y_continuous(expand = c(0, 0))
  labs(x = "Total length ()",
       y = "Weight ()")

weight_length_plot

```

## transform date column to class date, then find counts of observations by date
```{r}
lizard_counts_date <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(date)

lizard_counts_date

class(lizard_counts_date$date)

```
## make 
```{r}

date_plot <- ggplot(data = lizard_counts_date, aes(x = date, y = n)) +
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%Y") +
  labs(x = "Date",
       y = "Counts")

date_plot

```

```{r}

weight_length <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_stepsn(colors = c("green", "blue", "purple", "red", "orange"), 
                     breaks = c(0, 3, 5, 10, 30, 40))
  # scale_color_gradientn(colors = c("black", "green", "orange", "cyan", "purple"))
  # scale_color_gradient(low = "red", high = "blue")
  

weight_length

```

## update a color scheme using a palette in paletteer

### make a horizontal box plot with common_name (y), and total_length (x), with color changing based on common_name
```{r}
lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

name_length_plot <- ggplot(data = lizards_fct, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "rcartocolor::Earth")


name_length_plot
```


## 
```{r}

practice_plot <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    # panel.grid = element_blank()
    # plot.background = element_rect(fill = "cyan4")
    
    axis.text.x = element_text(color = "orange"),
    axis.title = element_text(color = "green",
                              size = 13),
    
    panel.background = element_rect(fill = "yellow",
                                    color = "purple",
                                    size = 15),
    panel.grid.major.x = element_line(color = "red"),
    panel.grid.major = element_line(color = "blue"),
    panel.grid.minor.y = element_line(color = "green")
    
    
  )
practice_plot
```

