---
title: "day9 Data Viz Continued"
author: "Steven Cognac"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# General use
library(tidyverse)
library(janitor)
library(here)

## specifically for data viz
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

library(sf)


```

## Get the lizards data
```{r}

lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

## ggrepel

## make subset

```{r}

ww_lizards <- lizards %>% 
  filter(common_name == "western whiptail",
         site == "sand")

ww_lizards_plot <- ggplot(data = ww_lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  labs(x = "Total length (cm)",
       y = "Weight (g)") +
  geom_text_repel(aes(label = toe_num),
                  size = 3,
                  max.overlaps = 10)

ww_lizards_plot

```

## Using gapminder, filter for only Europe, create a scatterplot with gdpPercap (x) and lifeExp (y), labeled by country name
```{r}


europe <- gapminder %>% 
  filter(continent == "Europe") %>% 
  filter(year == 1957)

europe_df <- data.frame(europe)

europe_gdp_1957 <- ggplot(data = europe, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country))

europe_gdp_1957

europe_gdp_1957 + 
```

## 

```{r}
p <- ggplot(data = lizards, aes(x = total_length, y = weight)) + 
  geom_point()

p + gghighlight(toe_num == 250, label_key = toe_num)
```
```{r}
q <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_line(aes(color = common_name))

## gghighlight lets you highlight things based on specific conditions in your dataset
q + gghighlight(max(weight) > 30)


```

## Patchwork library - for compound figures
 - makes it easier to put figures together in a single graphic
 - similar to facet_wrap(), but different

```{r}

# uses rules of PEMDAS to combine multiple plots & dataframes

(p | q) / q &
  theme_minimal()

```
# explore new graph types

### Marginal plots <- uses library(ggExtra)

```{r}
whiptail <- lizards %>% 
  filter(common_name== "western whiptail") %>% 
  drop_na()


```

```{r}
whiptail_plot <- ggplot(data = whiptail, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_rug()
  
  

whiptail_plot

```

## marginal plot with boxplots
```{r}

whiptail_plot <- ggplot(data = whiptail, aes(x = total_length, y = weight)) +
  geom_point(aes(color = sex), size = 2) +
  scale_color_manual(values = c("cyan4", "goldenrod", "black"),
                     name = "Sex:",
                     labels = c("female", "juvenile", "male")) +
  theme_minimal() +
  theme(
    
    legend.position = "bottom"
    
  )
                     

whiptail_plot

# create boxplot on margins of scatterplot
ggMarginal(whiptail_plot, type = "boxplot", groupColour = TRUE)

```
## Beeswarm plot

 - benefits of jitterplot & advantages of density or histogram
 - width of beeswarm plot is based on number of observations of the specific value
 
```{r}
swarm <- ggplot(data = whiptail, aes(x = sex, y = weight)) +
  geom_violin(fill = NA) +
  geom_beeswarm() +
  geom_boxplot(fill = NA)

swarm
```

## make a heatmap with geom_tile

### make new df named lizards counts with date converted to class Date. Then count by year and common_name # of lizards observed

```{r}

lizard_counts <- lizards %>% 
  
  ## create new date column and then get a year column
  mutate(date_new = lubridate::mdy(date)) %>% 
  mutate(year = lubridate::year(date_new)) %>% 
  
  # group by year and common_name
  # can also use count(year, common_name)
  
  group_by(year, common_name) %>% 
  summarize(counts = n())
  

```
## useful because you can show three variables on a two dimensional scale

```{r}
heat <- ggplot(data = lizard_counts, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = counts)) +
  geom_text(aes(label = counts),
            color = "white",
            size = 3) +
  scale_fill_viridis_c()

heat
            

```

## make a map of lizard observations
```{r}

jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml")) %>% 
  select(Name) %>% 
  clean_names()

map <- ggplot(data = jornada_veg) +
  geom_sf(aes(fill = name), color = NA, size = 0.01) +
   scale_fill_paletteer_d(palette = "ggthemes::manyeys") +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Dominant Vegetation") +
  theme_void()
  

map

ggsave(here("figs", "map.png"))

```

